Probability of causal contacts between interstellar civilizations through Monte Carlo simulations

### 2020-05-26 submit_v03/manuscript.tex
.

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Motivations}\label{S_motivations}
%{{{

@@

We stress the fact that the lack of detections of extraterrestrial intelligent signals
is
the only observational fact that can be taken

@@

Extraterrestrial Intelligences (SETI) have been working,
no signal has been received using the technical possibilities 
within their search criteria \citep{tarter_search_2001}.

@@

Notably, despite decades of effort not a single evidence for intelligent signals has been detected.

@@

In this work we address the problem of the temporal and spatial
structure of the distribution of communicating civilisations.
%
We make no asumptions about the origin of life, the development 
of intelligence or the formation of habitable planets according to stellar type.
%
Instead, we assume monoparametric function families to model the
appeareance of points in the disk of the Galaxy

%
To that end, we propose a model which depends on a set of only
three parameters.
%
We abandon the frequentist approach of the Drake equation to
compute the number of civilisations, providing its statistical
distribution instead.
%
This is an exploratory analysis that aims at developing a numerical
tool to discuss the different scenarios based on statistical
heuristics.
%
With this, we avoid the several theoretical problems
summarised by the Drake equation factors.
%
The approach proposed here should be considered as a compromise
between the uncertainties of the frequentist estimations and the detailed
recipes required on the numerical simulations.
%
We provide a framework to explore, through a suite of
numerical simulations, a parameter space of three unknown observables.
%
This allows to discuss possible scenarios and their consequences in
terms of the probability of making contacts.
%
The method we use for simulating a stochastic process is an
approximation that allows to study the behaviour of complex systems,
by considering a sequence of well defined discrete events.
%
The simulation is carried out by following all the variables that
describe and constitute the state of the system, and the evolution of
the process is described as a set of changes in that state.
%
In this context, an event produces a specific change in the state that
can be triggered by random variables that encode the stochastic nature
of the physical phenomenon.
%
For example, when a new contact is produced between two entities in
the simulated galaxy, the numbers of active communication lines and of
communicated nodes does change.
%
The process involves following the changes on the state of the system,
defining the initial and final states.
%
This is done by defining a method that allows to keep track of the
time progress in steps, and maintaining a list of relevant events.
%
In Sec.~\ref{S_methods} we introduce the methods and discuss the
candidate distributions for the statistical aspects of the times
involved in the communication process.
%
We present our results in Sec~\ref{S_results}, with special emphasis
on the statistical distributions of the duration of causal contacts in
one or both directions and the distribution of time intervals for the
waiting of the first contact.
%
This quantities are considered as a function of the three simulation
parameters.
%
In Sec.~\ref{S_discussion} we discuss our results and future research
directions.   
%
In this work we use the International System of units recommended by
the International Astronomial Union, where the simbol for years is
''a''.
%}}}

@@

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Methods and working hypotheses}\label{S_methods}
%{{{

@@

%{{{

>>>>

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Motivations}\label{S_motivations}

@@

We stress the fact that the lack of detections of extraterrestrial
intelligent signals is the only observational fact that can be taken

@@

Extraterrestrial Intelligences (SETI) have been working, no signal has
been received using the technical possibilities within their search
criteria \citep{tarter_search_2001}.

@@

Notably, despite decades of effort not a single evidence for
intelligent signals has been detected.

@@

In this work we address the problem of the temporal and spatial
structure of the distribution of communicating civilisations.
%
We make no asumptions about, e.g., the origin of life, the development
of intelligence or the formation of habitable planets according to
stellar type.
%
Instead, we assume monoparametric function families to model the
appearance of points in the disk of the Galaxy and over the time,
which we call \textit{nodes}.
%
These nodes represent the locations of ideal intelligent agents that
are able to receive and emit signals with perfect efficiency in all
directions.
%
Then we analyse the network of causally connected nodes limited to
have a maximum separation representing the maximum distance a signal
can travel with an intensity above a fixed threshold.
%
A node is causally connected to other node if it is within their light
cone.
%
We adopt as a definition that a light cone is the region in spacetime
within the surface generated by the light emanating from a given point
in space in a given period of time.
%
For example, two nodes separated by 100 light years which appear with
a difference in time of 80 years and last 50 years each, will be
causally connected for 30 years.
%
This connection is not bidirectional.
%
In fact, the second node will see the first one for 30 years, but the
first node will never acknowledge the existence of the second.
%
This simple tought experiment exemplify the importance of the time
variable to analyse the structure of connections.
%
The model can be described with three parameters, namely, the mean separation
between the appearance of nodes, the mean lifetime of the nodes, and
the maximum separation to allow an effective causal contact.
%
The model presented here comprises 
the stochastic networks of constrained causally connected nodes,
together with the three parameters and additional hypotheses.
%
Hereafter, we refeer to this model as the C3Net model, standing for
''Constrained Causally Conencted Network``.

We abandon the frequentist approach of the Drake equation to
compute the number of civilisations, providing its statistical
distribution instead.
%
This is an exploratory analysis that aims at developing a numerical
tool to discuss the different scenarios based on statistical
heuristics.
%
With this, we avoid the several theoretical problems
summarised by the Drake equation factors.
%
The approach proposed here should be considered as a compromise
between the uncertainties of the frequentist estimations and the detailed
recipes required on the numerical simulations.
%
We provide a framework to explore, through a suite of
numerical simulations, a parameter space of three unknown observables.
%
This allows to discuss possible scenarios and their consequences in
terms of the probability of making contacts.
%
The method we use for simulating a stochastic process is an
approximation that allows to study the behaviour of complex systems,
by considering a sequence of well defined discrete events.
%
The simulation is carried out by following all the variables that
describe and constitute the state of the system, and the evolution of
the process is described as a set of changes in that state.
%
In this context, an event produces a specific change in the state that
can be triggered by random variables that encode the stochastic nature
of the physical phenomenon.
%
For example, when a new contact is produced between two entities in
the simulated galaxy, the numbers of active communication lines and of
communicated nodes does change.
%
The process involves following the changes on the state of the system,
defining the initial and final states.
%
This is done by defining a method that allows to keep track of the
time progress in steps, and maintaining a list of relevant events.
%
In Sec.~\ref{S_methods} we introduce the methods and discuss the
candidate distributions for the statistical aspects of the times
involved in the communication process.
%
We present our results in Sec~\ref{S_results}, with special emphasis
on the statistical distributions of the duration of causal contacts in
one or both directions and the distribution of time intervals for the
waiting of the first contact.
%
This quantities are considered as a function of the three simulation
parameters.
%
In Sec.~\ref{S_discussion} we discuss our results and future research
directions.   
%
In this work we use the International System of units recommended by
the International Astronomial Union, where the simbol for years is
''a''.
%}}}

@@

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Methods and working hypotheses}\label{S_methods}

@@
